Visualizing probability distributions across bivariate cyclic temporal granularities

### 2020-03-11 main.Rmd
minor changes in quasi-periodic linear granularities and linear calendar algebra

If ${S_0,\dots,S_{R-1}}$ are the sets of indexes of $G$ describing
$H(0), \dots, H(R - 1)$, respectively, then the description of an arbitrary granule $H(j)$ is given by $\bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)$. Also, granularities can be periodical with respect to other granularities, except for a finite number of spans of time where they behave in an anomalous way [@Bettini2000-vy]. These are called quasi-periodic relationships. Intuitively, the definition requires that all granules of $H$ within the span of time between two exceptions have the same periodical behavior, characterized by $R$ and $P$.

@@

For example, in a Gregorian calendar without leap years each month is a grouping of the same number of days each year. The period is $P = 365$ and the number of granules of "months" in each period is given by $R=12$. Thus, we could say days group periodically into months. In a Gregorian calendar with leap years, days group quasi-periodically into months. The value of $P$ and $R$ would still be 365 and 12 respectively, while the <!--$E_i$s--> exceptions would be the time domain corresponding to $29^{\text{th}}$ February of any year.

@@

Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra [@Ning2002-tf]. It is assumed that there exists a *_bottom granularity_* and calendar algebra operations are designed to generate new granularities recursively from the bottom up. Some relevant calendar algebra operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} for illustrations in circular and aperiodic granularities.

@@

For example, due to the uneven length of months, we need to use the altering-tick operation to compute months from days.  Let us consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April (4), June (6), September (9) and November (11) by 1 day, and shrink February (2) by 3 days [@Ning2002-tf].

>>>>

If ${S_0,\dots,S_{R-1}}$ are the sets of indexes of $G$ describing
$H(0), \dots, H(R - 1)$, respectively, then the description of an arbitrary granule $H(j)$ is given by $\bigcup_{i \in S_j \mod R}G(P*\lfloor j/R \rfloor + i)$. Also, granularities can be periodical with respect to other granularities, except for a finite number of spans of time where they behave in an anomalous way [@Bettini2000-vy]. These are called quasi-periodic relationships.

<!-- CONTENT CUT -->
<!-- Intuitively, the definition requires that all granules of $H$ within the span of time between two exceptions have the same periodical behavior, characterized by $R$ and $P$. -->

@@

For example, in a Gregorian calendar without leap years we could say days group periodically into months with the period $P = 365$ and the number of granules of "months" in each period given by $R=12$. In a Gregorian calendar with leap years, days group quasi-periodically into months with the <!--The value of $P$ and $R$ would still be 365 and 12 respectively, while the $E_i$s--> exceptions of the time domain corresponding to $29^{\text{th}}$ February of any year.

@@

Linear time granularities are computed through an algebraic representation for time granularities, which is referred to as calendar algebra [@Ning2002-tf]. It is assumed that there exists a _bottom granularity_ and calendar algebra operations are designed to generate new granularities recursively from the bottom up. Some relevant calendar algebra operations are discussed below; these will be used in Section \ref{sec:cyclic-gran} for illustrations in circular and aperiodic cyclic granularities.

@@

For example, due to the uneven length of months, we need to use the altering-tick operation to compute months from days.  Let us consider a granularity "pseudomonth" which computes months only for non-leap years by grouping 31 days, and then shrink April, June, September and November by 1 day, and shrink February by 3 days [@Ning2002-tf].
